Evaluating infectious disease forecasts \\ with allocation scoring rules

### 2024-02-27 alloscore_manuscript/alloscore-application-shorter.Rnw
reorg to get all appendices referenced incrementally in main text.

We begin with an informal description of the allocation score and some examples illustrating its key characteristics in
section \ref{sec:methods.overview}. In section \ref{sec:methods.detailed} we develop the allocation score more
carefully, using a decision theoretic procedure for deriving proper scoring rules. (See appendix \ref{sec:proper}
for a definition of a proper scoring rule and a more technical discussion of the procedure). In section
\ref{sec:methods.related}, we note that another group of common scores including the quantile score, WIS, and CRPS, can
also be derived from decision theoretic foundations \textemdash starting from a different decision making context.

@@

\paragraph{Example 1} Suppose we have a forecast $F$ for need in two locations with $F_1 = \mathrm{Exp}(1 / \sigma_1)$
and $F_2 = \mathrm{Exp}(1 / \sigma_2)$, where $\sigma_1 = 1$ and $\sigma_2 = 4$. The means of these distributions are
given by the scale parameters $\sigma_i$. When the marginal forecasts are exponential distributions, it can be shown
that the optimal allocation divides the available resources among the locations proportionally to the scale parameters
$\sigma_i$ (see appendices \ref{sec:bayes-quantiles}). If $K = 5$ units of the resource are
available, the optimal allocation according to $F$ would be 1 and 4 resource units in locations 1 and 2, respectively. 
Increasing $K$ to $10$ changes this optimal allocation to 2 and 8. Figure~\ref{fig:exp_alloc_example} illustrates the 
situation.

@@

We arrive at a three-step procedure for developing scoring rules for probabilistic forecasts:
\begin{enumerate}
  \item Specify a \emph{loss function} $s(x, y)$ that measures the loss associated with taking action $x$ when outcome $y$
    eventually occurs.
  \item Given a probabilistic forecast $F$, determine the \emph{Bayes act} $x^F$ that minimizes the expected loss under
    the distribution $F$.
  \item The \emph{scoring rule} for $F$ calculates the score as the loss incurred when the Bayes act was used: 
    $S(F, y) = s(x^F, y)$.
\end{enumerate}
This is a general procedure that may be applied in settings where it is possible to specify a quantitative loss
function. We call a scoring rule obtained from this procedure a \emph{Bayes scoring rule} (noting that to our knowledge, 
this terminology is not standard).  In appendix \ref{sec:proper} we review the
result from the scoring rule literature (e.g., \cite{dawid2007geometry,gneiting2007strictly}) that Bayes scoring
rules are proper by construction.

@@

\paragraph{Step 2: Given a probabilistic forecast $F$, identify the Bayes act.} The Bayes act associated with the
forecast, $x^{F,K}$, is the allocation that minimizes the expected loss, that is, the solution of the \emph{allocation
problem} associated with $K$:
\begin{align}
  \underset{0 \leq x}{\mathrm{minimize}}\,\, \mathbb{E}_{F} [s_A(x, Y)] \text{ subject to }
  \, \sum_{i=1}^N x_i = K, \label{AP}
\end{align}
where $\mathbb{E}_{F} [s_A(x, Y)] = \sum_{i=1}^{N} L \cdot \mathbb{E}_{F_i}[\max(0, Y_i - x_i)]$ sums the expected loss
due to unmet need across all locations.

In appendix \ref{sec:bayes-quantiles} we show that the components of the Bayes act are quantiles 
$x_i^{F,K} = F_i^{-1}(\tau^{F,K})$ at a probability level $\tau^{F,K}$ that depends on the forecast $F$ and the resource
constraint $K$, but is shared across all locations. This probability level is the level at which the resource constraint
is satisfied: $\sum_{i=1}^N F_i^{-1}(\tau^{F,K}) = K$. This tells us that in order to allocate optimally (according to
$F$), we must divide resources among the locations so that there is an equal forecasted probability in every location
that the allocation is sufficient to meet resource need. This solution to the allocation problem is well-known in
inventory management and is often attributed to \cite{hadleywhitin1963}.

@@

As described above, predictions were submitted to the Forecast Hub in the form of a set of 23 quantiles of the
predictive distribution. The WIS can be directly calculated from these quantiles. However, calculation of the AS
requires that the full cumulative distribution functions of the forecast distributions for each location are available.
For the purpose of this analysis, to calculate the allocation score we approximated the full cumulative distribution
functions based on the provided quantiles. On the interior of the provided quantiles we used a monotonic cubic spline to
interpolate the quantiles, and in the lower and upper tails we used normal distributions with parameters selected so as
to match the two lowest and two highest quantiles (see appendix \ref{sec:distfromq} for further details).
In appendix \ref{sec:propriety_of_parametric_approximation}, we show that if this evaluation procedure
had been specified prospectively, the resulting score would be proper, \textemdash but that a post hoc application of this
procedure is improper. We use the procedure here to illustrate the properties of the AS, and note that a collaborative
forecasting hub interested in using the AS for evaluation could circumvent propriety issues by communicating the
definition of the AS to forecasters and then collecting allocations at specified resource levels as part of forecast
submissions.

@@

In practice, epidemiological forecasts are often directed at a diverse set of end-users. It may be easy for some of
these users to summarize the consequences of how they use a forecast with a numerical loss function, but for others,
such as officials deciding how best to update public ``situational awareness'' of an outbreak, this may be essentially
impossible. Even those forecast uses that are easily framed as expected loss minimization may differ enough that no
single score would be appropriate for all users. Ideally, targeted forecasting tools could be developed through close
collaboration between modelers and public health officials. However, this may only be possible in settings with
sufficient staffing on both an analytics and a public health team. Increasingly, collaborative modeling hubs are being
used to generate ``one-size-fits-all'' forecasts for many locations at once. But it could still be beneficial in these
settings to expand the set of scoring methods used by the hub to include scores, such as the AS, which are based on
specific public health decision problems.  This could help end users better understand the value of available forecasts
as inputs to their particular decision making contexts. (We note that issues of propriety are raised when forecasts are
evaluated with scoring rules for which they were not elicited or via parameters imputed by a hub after submission, both
of which are done in our application example.  We consider these issues briefly in appendices
\ref{sec:distfromq} and \ref{sec:propriety_of_parametric_approximation}.)

@@

\section{Introduction}
\label{sec:intro}

We address some technical and methodological points from the main text. We begin in section \ref{sec:proper} by defining
proper scoring rules and showing that the allocation score is proper. In section \ref{sec:ex-shortage}, we formalize the
concept of a \emph{shortage} of resources, give some key results about expected resource shortages under a distribution
characterizing uncertainty about (future) levels of resource need, and illustrate how quantiles arise as the Bayes act
in a decision making problem about the quantity of a resource to purchase. Section \ref{sec:bayes-quantiles} gives a
justification for the result that the Bayes act for the allocation problem is given by a vector of quantiles of the
forecast distributions in each location at a shared probability level, which was stated in section
\ref{sec:methods.detailed.specific_allocation}. We describe the algorithm that we use to compute allocations given a
forecast distribution in each location in section \ref{sec:numeric}. In section \ref{sec:distfromq}, we describe the
methods for approximating a full distribution from a set of quantiles, implemented in the R package \verb`distfromq`,
that we used to support computation of allocation scores from quantile forecasts that were submitted to the US COVID-19
Forecast Hub for the application in section \ref{sec:application}. Finally, in section
\ref{sec:propriety_of_parametric_approximation} we examine implications for propriety of an analysis that uses summaries
of forecasts (such as predictive quantiles) rather than the full forecast distributions for computation of allocation
scores.

\section{Proper scoring rules}
\label{sec:proper}

@@

\subsection{The allocation score is proper}
\label{sec:alloscore_proper}

@@

\section{Expected shortages}
\label{sec:ex-shortage}

A key feature of loss functions for decision problems used to define quantiles and related scoring rules such as the
CRPS and the WIS (see e.g., \cite{gneiting2011quantiles}, \cite{jose2009evaluating}, and \cite{royset2022optimization},
sections 1.C and 3.C), as well as the allocation loss function presented in this work, is the presence of a
\emph{shortage}: the amount $\max\{0,y-x\}$ by which a resource demand $y$ exceeds a supply decision variable $x$,
which, for convenience, we write as $(y-x)_{+}$. In particular, a quantile at probability level $\alpha$ of a
distribution $F$ on $\mathbb{R}^1$ (which we assume to have a well-defined density $f(x)$) is a Bayes act for the loss
function
\[
l(x,y) = Cx + L(y-x)_{+}
\]
where $\alpha = 1-C/L$ and $C$ and $L$ can be interpreted as the cost per unit of a resource (such as medicine) and the
loss incurred when a unit of demand (such as illness) cannot be met due to the shortage $(y-x)_{+}$.  This follows
because a Bayes act, as a minimizer of $\Ex_F[l(x,Y)]$, must also be a vanishing point of the derivative
\begin{align}
\dby{}{x} \Ex_F\left[l(x,Y)\right] &= \Ex_F\left[\dby{}{x}l(x,Y)\right] \nonumber\\
&= C + L\Ex_F\left[\dby{}{x}(Y-x)_+\right] \nonumber\\
&= C - L\Ex_F\left[\mathbf{1}\{Y > x\}\right] \nonumber\\
&= C + L(F(x) - 1), \label{eqn:q_deriv}
\end{align}
so that $1-C/L = F(x)$.
The formula $\dby{}{x}\Ex_F\left[(Y-x)_+\right] = F(x) - 1$ for the derivative of the shortage, used above in
\eqref{eqn:q_deriv}, can be obtained from an application of the ``Leibniz Rule'':
\begin{align}
	\frac{d}{dx} \Ex_F [(Y-x)_{+}] &= \frac{d}{dx} \int_{x}^{\infty} (y-x) f_Y(y)dy \nonumber\\
	&= \int_{x}^{\infty} \frac{d}{dx}(y-x) f_Y(y)dy - (x-x) f_Y(x) = -\int_{x}^{\infty} f_Y(y)dy = F(x)-1. \label{eqn:shortage_deriv}
\end{align}
Note that more care is required when $F$ does not have a density.
We will also use this result below in deriving the Bayes act for the allocation loss.

\section{Allocation Bayes acts as vectors of marginal quantiles.}
\label{sec:bayes-quantiles}

Here we study the form of the Bayes act for the allocation problem (AP) (equation \eqref{eqn:loss_fn} in section
\ref{sec:methods.detailed.specific_allocation}) of the text:
\begin{align}
    \underset{0 \leq x}{\mathrm{minimize}}\,\, \mathbb{E}_{F} [s_A(x, Y)]= \sum_{i=1}^{N} L \cdot \mathbb{E}_{F_i}[(Y_i - x_i)_{+}]
     \text{ subject to }
     \, \sum_{i=1}^N x_i = K, \label{AP-long}
\end{align}
where the marginal forecasts $F_i$ for $i=1,\dots,N$ represent forecasts for $N$ distinct locations. We show that the
Bayes act $x^{F,K} = (x_1^{F,K},\ldots,x_N^{F,K})$ for a forecast $F$ and resource constraint level $K$ is a vector of
quantiles of the marginal forecast distributions $F_i$ at a single probability level $\tau^{F,K}$, that is, $x_i^{F,K} =
q_{F_i,\tau^{F,K}}$. An immediate consequence used in the examples in Section \ref {sec:methods.overview} is that if
$F_i = \mathrm{Exp}(1/\sigma_i)$ for all $i$, then the Bayes act is proportional to $(\sigma_1,\ldots,\sigma_N)$, since
$q_{\mathrm{Exp}(1/\sigma),\tau} = -\sigma \log(1-\tau)$.

In order for $x^{\star} \in \mathbb{R}^N_{+}$ to solve the AP it must be true that reallocating $\delta > 0$ units of
the resource from location $i$ to location $j$ will lead to a net increase in expected shortage --- in other words, the
reallocation increases the expected shortage in location $i$ at least as much as it decreases the expected shortage
in location $j$:
\begin{align*}
&\mathbb{E}_{F_i}[(Y_i - (x^{\star}_i - \delta))_{+}] - \mathbb{E}_{F_i}[(Y_i - x^{\star}_i)_{+}]
\text{ (increase in $i$) } \nonumber \\
&\qquad \geq
\mathbb{E}_{F_j}[(Y_j - x^{\star}_j)_{+}] - \mathbb{E}_{F_j}[(Y_j - (x^{\star}_j + \delta))_{+}]
\text{ (decrease in $j$) }.
\end{align*}

Dividing by $\delta$ and letting $\delta \searrow 0$, this implies from
\eqref{eqn:shortage_deriv} that
\begin{align}
1-F_i(x^{\star}_i) &= -\frac{d}{dx_i}\mathbb{E}_{F_i}[(Y_i - x^{\star}_i)_{+}] \nonumber \\
&= \lim_{\delta \searrow 0} \frac{1}{\delta}
\left\{\mathbb{E}_{F_i}[(Y_i - (x^{\star}_i - \delta))_{+}] - \mathbb{E}_{F_i}[(Y_i - x^{\star}_i)_{+}]\right\}
\text{ (increase in $i$) } \nonumber \\
&\geq
\lim_{\delta \searrow 0} \frac{1}{\delta}
\left\{\mathbb{E}_{F_j}[(Y_j - x^{\star}_j)_{+}] - \mathbb{E}_{F_j}[(Y_j - (x^{\star}_j + \delta))_{+}]\right\}
\text{ (decrease in $j$) } \nonumber \\
 &= -\dby{}{x_j}\mathbb{E}_{F_j}[(Y_j - x^{\star}_j)_{+}] = 1-F_j(x^{\star}_j) \label{eqn:ASoptimal1}
\end{align}

Note that negative derivatives appear because our optimality condition addresses how a \emph{decrease} in resources will
\emph{increase} the expected shortage in $i$ and vice versa in $j$. Since \eqref{eqn:ASoptimal1} also holds with $i$ and $j$
reversed, a number $\lambda$ (a \emph{Lagrange multiplier}) exists such that
$L(1-F_k(x^{\star}_k)) = \lambda$ for all $k \in 1,\ldots,N$.
(We scale by $L$ to facilitate possible future interpretations of $\lambda$ in terms of the partial derivatives
of $\mathbb{E}_{F} [s_A(x, Y)]$.)
That is, $x^{\star}_k$ is a quantile $q_{\tau,F_k}$ for
$\tau = 1 - \lambda/L$. The value of $\tau$ is then determined by the constraint equation
\begin{align}
\sum_{i=1}^N q_{\tau,F_i} = K. \label{eqn:quantiles-sum-to-K}
\end{align}
It is important to note that $\tau$ depends on $F$ and $K$ and is \emph{not} a fixed parameter
of the allocation scoring rule.

\section{Numerical computation of allocation Bayes acts}
\label{sec:numeric}

To compute an allocation score $S_A(F,y;K) := s_A(x^{F,K},y)$, we require numerical values for a Bayes act solving the
AP \eqref{AP-long} --- that is, we must find the specific resource allocations for each location that are determined by
the forecast $F$ under the resource constraint $K$. Assuming we have reliable means of calculating quantiles
$q_{\alpha,F_i}$ of the marginal forecasts $F_i$, these allocations are given by $q_{\tau^{\star},F_i}$ where
$\tau^{\star}$ solves the equation \eqref{eqn:quantiles-sum-to-K}. However, this equation is not analytically tractable
and we must resort to a numerical method for finding an approximation $\tilde{\tau}$ of $\tau^{\star}$.

@@

\section{Computing allocations from finite quantile forecast representations}
\label{sec:distfromq}

@@

To address this situation and support the numerical allocation algorithm outlined in section \ref{sec:numeric}, we need
a mechanism to approximate the full cumulative distribution functions $F_i$, $i = 1, \ldots, N$ based on the provided
quantiles. We have developed functionality for this purpose in the \verb`distfromq` package for R \citep{ray-distfromq}.
This functionality represents a distribution as a mixture of discrete and continuous parts, and it works in two steps:
\begin{enumerate}
  \item Identify a discrete component of the distribution consisting of zero or more point masses, and create an
    adjusted set of predictive quantiles for the continuous part of the distribution by subtracting the point mass
    probabilities and rescaling.
  \item For the continuous part of the distribution, different approaches are used on the interior and exterior of the
    provided quantiles:
  \begin{enumerate}
    \item On the interior, a monotonic cubic spline interpolates the adjusted quantiles representing the continuous part
      of the distribution.
    \item A location-scale parametric family is used to extrapolate beyond the provided quantiles. The location and
      scale parameters are estimated separately for the lower and upper tails so as to obtain a tail distribution that
      matches the two most extreme quantiles in each tail. In this work, we use normal distributions for the tails.
  \end{enumerate}
\end{enumerate}
The resulting distributional estimate exactly matches all of the predictive quantiles provided by the forecaster. We use
the cumulative distribution function resulting from this procedure as an input to the allocation score algorithm.

@@

\section{Propriety of parametric approximation} % (fold)
\label{sec:propriety_of_parametric_approximation}

@@

We then arrive at the scoring rule
$$S_R(F, y) = s_R(\theta^\star(F), y) = s_A(x^{G_{\theta^\star(F)}}, y).$$
It follows from the discussion in section \ref{sec:alloscore_proper} that this is a proper scoring rule for $F$.
Although the full forecast distribution $F$ is not available in the forecast submission, the score $S_R(F, y)$ can be
calculated from the reported parameter values as long as the forecaster submits the optimal parameters
$\theta^\star(F)$.

@@

We define
$$G^{\star}(F) := \argmin_{G \in \mathcal{P}_{\mathrm{dfq}}} E_{F}[S_A(G, Y)].$$
Since $S_R$ is defined as the Bayes scoring rule for the forecaster's loss \eqref{eqn:forecaster_theta_loss}, $G^{\star}(F)$ coincides
with $G_{\theta^{\star}(F)}$, the distribution in $\mathcal{P}_{\mathrm{dfq}}$ given by the
optimal submission parameters $\theta^{\star}(F)$ for the forecaster with predictive distribution $F$.
In general, $G^q(F)$ and $G^\star(F)$ will be different distributions:  matching $F$ at specific quantiles
does not require $G^q(F)$ to match $F$ at the quantiles for $\tau^{F,K}$ (c.f. \eqref{eqn:quantiles-sum-to-K}),
which would be necessary for it to share $x^F$ as an optimal allocation.

>>>>

We begin with an informal description of the allocation score and some examples illustrating its key characteristics in
section \ref{sec:methods.overview}. In section \ref{sec:methods.detailed} we develop the allocation score more
carefully, using a decision theoretic procedure for deriving proper scoring rules. (See appendix \ref{sec:a:proper}
for a definition of a proper scoring rule and a more technical discussion of the procedure). In section
\ref{sec:methods.related}, we note that another group of common scores including the quantile score, WIS, and CRPS, can
also be derived from decision theoretic foundations \textemdash starting from a different decision making context.

@@

\paragraph{Example 1} Suppose we have a forecast $F$ for need in two locations with $F_1 = \mathrm{Exp}(1 / \sigma_1)$
and $F_2 = \mathrm{Exp}(1 / \sigma_2)$, where $\sigma_1 = 1$ and $\sigma_2 = 4$. The means of these distributions are
given by the scale parameters $\sigma_i$. When the marginal forecasts are exponential distributions, it can be shown
that the optimal allocation divides the available resources among the locations proportionally to the scale parameters
$\sigma_i$ (see appendix \ref{sec:a:bayes-quantiles}). If $K = 5$ units of the resource are
available, the optimal allocation according to $F$ would be 1 and 4 resource units in locations 1 and 2, respectively. 
Increasing $K$ to $10$ changes this optimal allocation to 2 and 8. Figure~\ref{fig:exp_alloc_example} illustrates the 
situation.

@@

We arrive at a three-step procedure for developing scoring rules for probabilistic forecasts:
\begin{enumerate}
  \item Specify a \emph{loss function} $s(x, y)$ that measures the loss associated with taking action $x$ when outcome $y$
    eventually occurs.
  \item Given a probabilistic forecast $F$, determine the \emph{Bayes act} $x^F$ that minimizes the expected loss under
    the distribution $F$.
  \item The \emph{scoring rule} for $F$ calculates the score as the loss incurred when the Bayes act was used: 
    $S(F, y) = s(x^F, y)$.
\end{enumerate}
This is a general procedure that may be applied in settings where it is possible to specify a quantitative loss
function. We call a scoring rule obtained from this procedure a \emph{Bayes scoring rule} (noting that to our knowledge, 
this terminology is not standard).  In appendix \ref{sec:a:proper} we review the
result from the scoring rule literature (e.g., \cite{dawid2007geometry,gneiting2007strictly}) that Bayes scoring
rules are proper by construction.

@@

\paragraph{Step 2: Given a probabilistic forecast $F$, identify the Bayes act.} The Bayes act associated with the
forecast, $x^{F,K}$, is the allocation that minimizes the expected loss, that is, the solution of the \emph{allocation
problem} associated with $K$:
\begin{align}
  \underset{0 \leq x}{\mathrm{minimize}}\,\, \mathbb{E}_{F} [s_A(x, Y)] \text{ subject to }
  \, \sum_{i=1}^N x_i = K, \label{eqn:AP}
\end{align}
where $\mathbb{E}_{F} [s_A(x, Y)] = \sum_{i=1}^{N} L \cdot \mathbb{E}_{F_i}[\max(0, Y_i - x_i)]$ sums the expected loss
due to unmet need across all locations.

In appendix \ref{sec:a:bayes-quantiles} we show that the components of the Bayes act are quantiles 
$x_i^{F,K} = F_i^{-1}(\tau^{F,K})$ at a probability level $\tau^{F,K}$ that depends on the forecast $F$ and the resource
constraint $K$, but is shared across all locations. This probability level is the level at which the resource constraint
is satisfied: $\sum_{i=1}^N F_i^{-1}(\tau^{F,K}) = K$. This tells us that in order to allocate optimally (according to
$F$), we must divide resources among the locations so that there is an equal forecasted probability in every location
that the allocation is sufficient to meet resource need. This solution to the allocation problem is well-known in
inventory management and is often attributed to \cite{hadleywhitin1963}.

@@

As described above, predictions were submitted to the Forecast Hub in the form of a set of 23 quantiles of the
predictive distribution. The WIS can be directly calculated from these quantiles. However, our numerical method for calculating 
the AS (outlined in appendix \ref{sec:a:numeric})
requires a full cumulative distribution function (CDF) for the forecast in each location.
For the purpose of this analysis, we imputed CDFs based on the provided quantiles following a procedure detailed in 
appendix \ref{sec:a:distfromq}. Briefly, this involves interpolating the provided quantiles between the lowest (.01) and highest 
(.99) probability levels using monotonic cubic splinea, and then extending outside of these level with normal distribution tails
parametrized to match the two lowest and two highest quantiles. In appendix \ref{sec:a:propriety_of_parametric_approximation}, 
we show that if this evaluation procedure
had been specified prospectively, the resulting score would be proper, \textemdash but that a post hoc application of this
procedure is improper. We use the procedure here to illustrate the properties of the AS, and note that a collaborative
forecasting hub interested in using the AS for evaluation could circumvent propriety issues by communicating the
definition of the AS to forecasters and then collecting allocations at specified resource levels as part of forecast
submissions.

@@

In practice, epidemiological forecasts are often directed at a diverse set of end-users. It may be easy for some of
these users to summarize the consequences of how they use a forecast with a numerical loss function, but for others,
such as officials deciding how best to update public ``situational awareness'' of an outbreak, this may be essentially
impossible. Even those forecast uses that are easily framed as expected loss minimization may differ enough that no
single score would be appropriate for all users. Ideally, targeted forecasting tools could be developed through close
collaboration between modelers and public health officials. However, this may only be possible in settings with
sufficient staffing on both an analytics and a public health team. Increasingly, collaborative modeling hubs are being
used to generate ``one-size-fits-all'' forecasts for many locations at once. But it could still be beneficial in these
settings to expand the set of scoring methods used by the hub to include scores, such as the AS, which are based on
specific public health decision problems.  This could help end users better understand the value of available forecasts
as inputs to their particular decision making contexts. (We note that issues of propriety are raised when forecasts are
evaluated with scoring rules for which they were not elicited or via parameters imputed by a hub after submission, both
of which are done in our application example.  We consider these issues briefly in appendices
\ref{sec:a:distfromq} and \ref{sec:a:propriety_of_parametric_approximation}.)

@@

We address some technical and methodological points from the main text. We begin in section \ref{sec:a:proper} by defining
proper scoring rules and showing that the allocation score is proper. Section \ref{sec:a:bayes-quantiles} gives a
justification for the result that the Bayes act for the allocation problem is given by a vector of quantiles of the
forecast distributions in each location at a shared probability level, which was stated in section
\ref{sec:methods.detailed.specific_allocation}. We describe the algorithm that we use to compute allocations given a
forecast distribution in each location in section \ref{sec:a:numeric}. In section \ref{sec:a:distfromq}, we describe the
methods for approximating a full distribution from a set of quantiles, implemented in the R package \verb`distfromq`,
that we used to support computation of allocation scores from quantile forecasts that were submitted to the US COVID-19
Forecast Hub for the application in section \ref{sec:application}. Finally, in section
\ref{sec:a:propriety_of_parametric_approximation} we examine implications for propriety of an analysis that uses summaries
of forecasts (such as predictive quantiles) rather than the full forecast distributions for computation of allocation
scores.

\section{Proper scoring rules}
\label{sec:a:proper}

@@

\subsection{The allocation score is proper}
\label{sec:a:alloscore_proper}

@@

\section{Allocation Bayes acts as vectors of marginal quantiles.}
\label{sec:a:bayes-quantiles}

Here we study the form of the Bayes act for the allocation problem (AP) (equation \eqref{eqn:AP} in section
\ref{sec:methods.detailed.specific_allocation}) of the text:
\begin{align}
    \underset{0 \leq x}{\mathrm{minimize}}\,\, 
    \mathbb{E}_{F} [s_A(x, Y)]= \sum_{i=1}^{N} L \cdot \mathbb{E}_{F_i}[\max(0, Y_i - x_i)]
     \text{ subject to }
     \, \sum_{i=1}^N x_i = K, \label{eqn:AP-long}
\end{align}
where the marginal forecasts $F_i$ for $i=1,\dots,N$ represent forecasts for $N$ distinct locations. We show that the
Bayes act $x^{F,K} = (x_1^{F,K},\ldots,x_N^{F,K})$ for a forecast $F$ and resource constraint level $K$ is a vector of
quantiles of the marginal forecast distributions $F_i$ at a single probability level $\tau^{F,K}$, that is, $x_i^{F,K} =
q_{F_i,\tau^{F,K}}$. An immediate consequence used in the examples in Section \ref {sec:methods.overview} is that if
$F_i = \mathrm{Exp}(1/\sigma_i)$ for all $i$, then the Bayes act is proportional to $(\sigma_1,\ldots,\sigma_N)$, since
$q_{\mathrm{Exp}(1/\sigma),\tau} = -\sigma \log(1-\tau)$.

We begin by noting that a key feature of each term of the loss function $s_A(x, Y)$ defining the AP is the presence of 
a \emph{shortage}: an amount $\max\{0,y-x\}$ by which a resource demand $y$ exceeds a supply decision variable $x$,
which, for convenience, we write as $(y-x)_{+}$.  This is a feature shared with decision problems used to define quantiles 
and related scoring rules such as the
CRPS and the WIS (see e.g., \cite{gneiting2011quantiles}, \cite{jose2009evaluating}, and \cite{royset2022optimization},
sections 1.C and 3.C). In particular, 
a quantile at probability level $\alpha$ of a
distribution $F$ on $\mathbb{R}$ (which we assume to have a well-defined density $f(x)$) is a Bayes act for the loss
function
\[
l(x,y) = Cx + L(y-x)_{+}
\]
where $\alpha = 1-C/L$ and $C$ and $L$ can be interpreted as the cost per unit of a resource (such as medicine) and the
loss incurred when a unit of demand (such as illness) cannot be met due to the shortage $(y-x)_{+}$.  This follows
because a Bayes act, as a minimizer of $\Ex_F[l(x,Y)]$, must also be a vanishing point of the derivative
\begin{align}
\dby{}{x} \Ex_F\left[l(x,Y)\right] &= \Ex_F\left[\dby{}{x}l(x,Y)\right] \nonumber\\
&= C + L\Ex_F\left[\dby{}{x}(Y-x)_+\right] \nonumber\\
&= C - L\Ex_F\left[\mathbf{1}\{Y > x\}\right] \nonumber\\
&= C + L(F(x) - 1), \label{eqn:q_deriv}
\end{align}
so that $1-C/L = F(x)$.
The formula $\dby{}{x}\Ex_F\left[(Y-x)_+\right] = F(x) - 1$ for the derivative of the shortage, used above in
\eqref{eqn:q_deriv}, can be obtained from an application of the ``Leibniz Rule'':
\begin{align}
	\frac{d}{dx} \Ex_F [(Y-x)_{+}] &= \frac{d}{dx} \int_{x}^{\infty} (y-x) f_Y(y)dy \nonumber\\
	&= \int_{x}^{\infty} \frac{d}{dx}(y-x) f_Y(y)dy - (x-x) f_Y(x) = -\int_{x}^{\infty} f_Y(y)dy = F(x)-1. \label{eqn:shortage_deriv}
\end{align}
Note that more care is required when $F$ does not have a density.

Returning to the AP \eqref{eqn:AP-long}, notice that in order for $x^{\star} \in \mathbb{R}^N_{+}$ to be a Bayes act it must be true that reallocating $\delta > 0$ units of
the resource from location $i$ to location $j$ will lead to a net increase in expected shortage --- in other words, the
reallocation increases the expected shortage in location $i$ at least as much as it decreases the expected shortage
in location $j$:
\begin{align*}
&\mathbb{E}_{F_i}[(Y_i - (x^{\star}_i - \delta))_{+}] - \mathbb{E}_{F_i}[(Y_i - x^{\star}_i)_{+}]
\text{ (detriment in $i$) } \nonumber \\
&\qquad \geq
\mathbb{E}_{F_j}[(Y_j - x^{\star}_j)_{+}] - \mathbb{E}_{F_j}[(Y_j - (x^{\star}_j + \delta))_{+}]
\text{ (benefit in $j$) }.
\end{align*}

Dividing by $\delta$ and letting $\delta \searrow 0$, this implies from
\eqref{eqn:shortage_deriv} that
\begin{align}
1-F_i(x^{\star}_i) &= -\frac{d}{dx_i}\mathbb{E}_{F_i}[(Y_i - x^{\star}_i)_{+}] 
\text{ (rate of detriment in $i$) }\nonumber \\
&= \lim_{\delta \searrow 0} \frac{1}{\delta}
\left\{\mathbb{E}_{F_i}[(Y_i - (x^{\star}_i - \delta))_{+}] - \mathbb{E}_{F_i}[(Y_i - x^{\star}_i)_{+}]\right\}
\nonumber \\
&\geq
\lim_{\delta \searrow 0} \frac{1}{\delta}
\left\{\mathbb{E}_{F_j}[(Y_j - x^{\star}_j)_{+}] - \mathbb{E}_{F_j}[(Y_j - (x^{\star}_j + \delta))_{+}]\right\}
 \nonumber \\
 &= -\dby{}{x_j}\mathbb{E}_{F_j}[(Y_j - x^{\star}_j)_{+}] = 1-F_j(x^{\star}_j) 
 \text{ (rate of benefit in $j$).}\label{eqn:ASoptimal1}
\end{align}
(Negative derivatives appear here because our optimality condition addresses how a \emph{decrease} in resources will
 \emph{increase} the expected shortage in $i$ and vice versa in $j$.)

Since \eqref{eqn:ASoptimal1} also holds with $i$ and $j$
reversed, a number $\lambda$ (a \emph{Lagrange multiplier}) exists such that
$L(1-F_k(x^{\star}_k)) = \lambda$ for all $k \in 1,\ldots,N$.
(We scale by $L$ to facilitate possible future interpretations of $\lambda$ in terms of the partial derivatives
of $\mathbb{E}_{F} [s_A(x, Y)]$.)
That is, $x^{\star}_k$ is a quantile $q_{\tau,F_k}$ for
$\tau = 1 - \lambda/L$. The value of $\tau$ is then determined by the constraint equation
\begin{align}
\sum_{i=1}^N q_{\tau,F_i} = K. \label{eqn:quantiles-sum-to-K}
\end{align}
It is important to note that $\tau$ depends on $F$ and $K$ and is \emph{not} a fixed parameter
of the allocation scoring rule.

\section{Numerical computation of allocation Bayes acts}
\label{sec:a:numeric}

To compute an allocation score $S_A(F,y;K) := s_A(x^{F,K},y)$, we require numerical values for a Bayes act solving the
AP \eqref{eqn:AP-long} --- that is, we must find the specific resource allocations for each location that are determined by
the forecast $F$ under the resource constraint $K$. Assuming we have reliable means of calculating quantiles
$q_{\alpha,F_i}$ of the marginal forecasts $F_i$, these allocations are given by $q_{\tau^{\star},F_i}$ where
$\tau^{\star}$ solves the equation \eqref{eqn:quantiles-sum-to-K}. However, this equation is not analytically tractable
and we must resort to a numerical method for finding an approximation $\tilde{\tau}$ of $\tau^{\star}$.

@@

\section{Computing allocations from finite quantile forecast representations}
\label{sec:a:distfromq}

@@

To address this situation and support the numerical allocation algorithm outlined in section \ref{sec:a:numeric}, we need
a mechanism to approximate the full cumulative distribution functions $F_i$, $i = 1, \ldots, N$ based on the provided
quantiles. We have developed functionality for this purpose in the \verb`distfromq` package for R \citep{ray-distfromq}.
This functionality represents a distribution as a mixture of discrete and continuous parts, and it works in two steps:
\begin{enumerate}
  \item Identify a discrete component of the distribution consisting of zero or more point masses, and create an
    adjusted set of predictive quantiles for the continuous part of the distribution by subtracting the point mass
    probabilities and rescaling.
  \item For the continuous part of the distribution, different approaches are used on the interior and exterior of the
    provided quantiles:
  \begin{enumerate}
    \item On the interior, a monotonic cubic spline interpolates the adjusted quantiles representing the continuous part
      of the distribution.
    \item A location-scale parametric family is used to extrapolate beyond the provided quantiles. The location and
      scale parameters are estimated separately for the lower and upper tails so as to obtain a tail distribution that
      matches the two most extreme quantiles in each tail. In this work, we use normal distributions for the tails.
  \end{enumerate}
\end{enumerate}
The resulting distributional estimate exactly matches all of the predictive quantiles provided by the forecaster. We use
the cumulative distribution function resulting from this procedure as an input to the allocation score algorithm.

@@

\section{Propriety of parametric approximation} % (fold)
\label{sec:a:propriety_of_parametric_approximation}

@@

We then arrive at the scoring rule
$$S_R(F, y) = s_R(\theta^\star(F), y) = s_A(x^{G_{\theta^\star(F)}}, y).$$
It follows from the discussion in section \ref{sec:a:alloscore_proper} that this is a proper scoring rule for $F$.
Although the full forecast distribution $F$ is not available in the forecast submission, the score $S_R(F, y)$ can be
calculated from the reported parameter values as long as the forecaster submits the optimal parameters
$\theta^\star(F)$.

@@

We define
$$G^{\star}(F) := \argmin_{G \in \mathcal{P}_{sec:propriety_of_parametric_approximation\mathrm{dfq}}} E_{F}[S_A(G, Y)].$$
Since $S_R$ is defined as the Bayes scoring rule for the forecaster's loss \eqref{eqn:forecaster_theta_loss}, $G^{\star}(F)$ coincides
with $G_{\theta^{\star}(F)}$, the distribution in $\mathcal{P}_{\mathrm{dfq}}$ given by the
optimal submission parameters $\theta^{\star}(F)$ for the forecaster with predictive distribution $F$.
In general, $G^q(F)$ and $G^\star(F)$ will be different distributions:  matching $F$ at specific quantiles
does not require $G^q(F)$ to match $F$ at the quantiles for $\tau^{F,K}$ (c.f. \eqref{eqn:quantiles-sum-to-K}),
which would be necessary for it to share $x^F$ as an optimal allocation.
